---
title: "Vietnamese Healthcare Cost Prediction — Improved"
author: "q"
date: "2026-05-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## 1. Libraries

```{r library}
library(readxl)
library(tidyverse)
library(readr)
library(fastDummies)
library(xgboost)
library(caret)
```

---

## 2. Import Data

```{r import}
train_raw <- read_csv("train.csv")
test_raw  <- read_csv("test.csv")
```

---

## 3. Clean Data

```{r clean}
CAT_COLS <- c(
  "GENDER", "MARRIED_STATUS",
  "LOAI_THE_1", "LOAI_THE_2",
  "CO_LAM_VIEC", "CO_DUOC_KY_HOP_DONG",
  "CO_CONG_VIEC_THU_HAI", "CO_DUOC_KY_HOP_DONG_2",
  "MA_NGHE_NGHIEP", "MA_NGHE_2",
  "AGE_GROUP"
)

# Columns that exist only in train — used in Stage 1 stacking
TARGET_COLS <- c("FRE_NGOAITRU", "FRE_NOITRU", "COST_NGOAITRU", "COST_NOITRU")

clean_data <- function(df) {
  df %>%
    select(-ID) %>%

    # Flag newborns
    mutate(AGE_IS_ZERO = as.integer(AGE == 0)) %>%

    # Structural NAs: no job → income = 0
    mutate(across(
      c(TIEN_LUONG_12T, TIEN_LE_TET, THUONG_PHU_CAP,
        LUONG_CONG_VIEC_2, TIEN_LE_TET_2, THUONG_PHU_CAP_2,
        LUONG_CONG_VIEC_3,
        TIEN_TRO_CAP_THAT_NGHIEP, TIEN_TRO_CAP_THOI_VIEC_1_LAN,
        LUONG_CHO_NGHI_HUU_BINH_THUONG, LUONG_CHO_NGHI_HUU_SOM,
        TRO_CAP_MAT_SUC_LAO_DONG),
      ~ replace_na(., 0)
    )) %>%

    # Categorical NAs → "UNKNOWN" / "NONE"
    mutate(
      MARRIED_STATUS        = replace_na(as.character(MARRIED_STATUS), "UNKNOWN"),
      LOAI_THE_1            = replace_na(as.character(LOAI_THE_1),     "UNKNOWN"),
      LOAI_THE_2            = replace_na(as.character(LOAI_THE_2),     "UNKNOWN"),
      CO_LAM_VIEC           = replace_na(as.character(CO_LAM_VIEC),    "UNKNOWN"),
      CO_DUOC_KY_HOP_DONG   = replace_na(as.character(CO_DUOC_KY_HOP_DONG),   "UNKNOWN"),
      CO_DUOC_KY_HOP_DONG_2 = replace_na(as.character(CO_DUOC_KY_HOP_DONG_2), "UNKNOWN"),
      MA_NGHE_2             = replace_na(as.character(MA_NGHE_2), "NONE"),
      GENDER                = as.character(GENDER),
      CO_CONG_VIEC_THU_HAI  = as.character(CO_CONG_VIEC_THU_HAI)
    ) %>%

    # Education: median impute (~14 % missing)
    mutate(
      HOC_DEN_LOP_MAY = replace_na(HOC_DEN_LOP_MAY,
                                    median(HOC_DEN_LOP_MAY, na.rm = TRUE))
    )
}

train_clean <- clean_data(train_raw)
test_clean  <- clean_data(test_raw)
```

---

## 4. Feature Engineering

```{r feature-engineering}
fe <- function(df, is_train = FALSE) {
  # ── Base features ─────────────────────────────────────────────────────────
  out <- df %>%
    mutate(
      TOTAL_INCOME = TIEN_LUONG_12T + TIEN_LE_TET + THUONG_PHU_CAP +
                     LUONG_CONG_VIEC_2 + TIEN_LE_TET_2 + THUONG_PHU_CAP_2 +
                     LUONG_CONG_VIEC_3,

      HAS_INCOME_2 = as.integer(LUONG_CONG_VIEC_2 > 0),

      AGE_GROUP = case_when(
        AGE < 6   ~ "infant",
        AGE < 18  ~ "child",
        AGE < 40  ~ "young_adult",
        AGE < 60  ~ "middle_age",
        TRUE      ~ "elderly"
      ),

      AGE_SQ     = AGE ^ 2,
      IS_ELDERLY = as.integer(AGE >= 60),

      TOTAL_SOCIAL = TIEN_TRO_CAP_THAT_NGHIEP + TIEN_TRO_CAP_THOI_VIEC_1_LAN +
                     LUONG_CHO_NGHI_HUU_BINH_THUONG + LUONG_CHO_NGHI_HUU_SOM +
                     TRO_CAP_MAT_SUC_LAO_DONG,

      HAS_SOCIAL_SUPPORT = as.integer(TOTAL_SOCIAL > 0)
    )

  # ── 11.4 Interaction & ratio features ─────────────────────────────────────
  out <- out %>%
    mutate(
      # Income per year of age — proxy for socioeconomic status
      INCOME_PER_AGE = TOTAL_INCOME / (AGE + 1),

      # Log-income — compress extreme earners
      LOG_INCOME     = log1p(TOTAL_INCOME),

      # Log-social income
      LOG_SOCIAL     = log1p(TOTAL_SOCIAL),

      # Insurance card interaction (use raw columns before OHE)
      HAS_BOTH_CARDS = as.integer(LOAI_THE_1 != "UNKNOWN" & LOAI_THE_2 != "UNKNOWN"),

      # Age × income interaction
      AGE_X_INCOME   = AGE * log1p(TOTAL_INCOME)
    )

  out
}

train_fe <- fe(train_clean, is_train = TRUE)
test_fe  <- fe(test_clean,  is_train = FALSE)

cat("Cols in train but not in test (should be target cols only):\n")
print(setdiff(names(train_fe), names(test_fe)))
```

---

## 5. Outlier Capping (11.7)

Cap numeric features at the 99th percentile to reduce influence of extreme values.
The cap thresholds are fitted on train only and applied to test.

```{r outlier-cap}
num_cols <- c("AGE", "TIEN_LUONG_12T", "TIEN_LE_TET", "THUONG_PHU_CAP",
              "LUONG_CONG_VIEC_2", "TOTAL_INCOME", "TOTAL_SOCIAL",
              "INCOME_PER_AGE", "LOG_INCOME", "LOG_SOCIAL")

# Fit caps on train
cap_vals <- train_fe %>%
  summarise(across(all_of(num_cols), ~ quantile(., 0.99, na.rm = TRUE))) %>%
  as.list()

cap_col <- function(df, col, cap) {
  df[[col]] <- pmin(df[[col]], cap)
  df
}

for (col in num_cols) {
  train_fe <- cap_col(train_fe, col, cap_vals[[col]])
  test_fe  <- cap_col(test_fe,  col, cap_vals[[col]])
}

# Boxplot after capping
train_fe %>%
  select(all_of(num_cols)) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x = name, y = value)) +
  geom_boxplot(outlier.colour = "red", outlier.size = 0.5) +
  facet_wrap(~name, scales = "free") +
  theme_minimal() +
  labs(title = "Boxplot of numeric features (after 99th-pct capping)")
```

---

## 6. One-Hot Encoding & Column Alignment

```{r dummy}
encode <- function(df, cat_cols) {
  dummy_cols(df,
             select_columns          = cat_cols,
             remove_selected_columns = TRUE,
             remove_first_dummy      = TRUE)
}

train_dummy <- encode(train_fe, CAT_COLS)
test_dummy  <- encode(test_fe,  CAT_COLS)

# Feature columns = train cols minus target and label
train_feature_cols <- setdiff(names(train_dummy), c("y", TARGET_COLS))

# Add any columns that are in train but missing from test (unseen levels)
missing_cols <- setdiff(train_feature_cols, names(test_dummy))
if (length(missing_cols) > 0) {
  cat("Adding missing columns to test:", missing_cols, "\n")
  test_dummy[missing_cols] <- 0
}

# Keep only the training feature columns, in the same order
test_dummy <- test_dummy[, train_feature_cols]

cat("Train:", nrow(train_dummy), "×", length(train_feature_cols), "\n")
cat("Test :", nrow(test_dummy),  "×", ncol(test_dummy),   "\n")
cat("Column match:", all(train_feature_cols == names(test_dummy)), "\n")
```

---

## 7. Prepare Matrices

```{r prep-matrices}
X_train <- train_dummy[, train_feature_cols] %>% as.data.frame()
y_train <- train_dummy$y
X_test  <- test_dummy %>% as.data.frame()

# Sanity check — no data leak
leak_check <- intersect(names(X_train),
                        c("FRE_NGOAITRU", "FRE_NOITRU",
                          "COST_NGOAITRU", "COST_NOITRU", "y"))
if (length(leak_check) == 0) cat("OK — no data leak\n") else
  cat("WARNING data leak:", leak_check, "\n")
```

---

## 8. Stage 1 — Out-of-Fold Sub-component Predictions (11.1 Stacking)

Build out-of-fold (OOF) predictions for each of the 4 sub-components using 5-fold CV.
These OOF predictions become extra features for the final model — no leakage because
each row is predicted by a model that never saw that row during training.
Test predictions are the mean of 5 fold models.

```{r oof-stacking}
set.seed(42)
K       <- 5
folds   <- createFolds(y_train, k = K, list = TRUE, returnTrain = FALSE)

# Quick xgboost params for sub-component models
base_params <- list(
  objective        = "reg:squarederror",
  max_depth        = 5,
  eta              = 0.05,
  subsample        = 0.8,
  colsample_bytree = 0.8,
  min_child_weight = 10,
  nthread          = parallel::detectCores()
)

# We train sub-models on log1p scale as well (right-skewed counts/costs)
oof_predict_component <- function(component_name, X, y_comp, X_test_mat) {
  oof_preds <- numeric(nrow(X))
  test_preds_list <- vector("list", K)
  y_log <- log1p(y_comp)

  for (k in seq_len(K)) {
    val_idx   <- folds[[k]]
    train_idx <- setdiff(seq_len(nrow(X)), val_idx)

    dtrain <- xgb.DMatrix(as.matrix(X[train_idx, ]), label = y_log[train_idx])
    dval   <- xgb.DMatrix(as.matrix(X[val_idx,   ]), label = y_log[val_idx])
    dtest  <- xgb.DMatrix(X_test_mat)

    m <- xgb.train(
      params  = base_params,
      data    = dtrain,
      nrounds = 500,
      watchlist = list(val = dval),
      early_stopping_rounds = 30,
      verbose = 0
    )

    oof_preds[val_idx]  <- expm1(predict(m, dval))
    test_preds_list[[k]] <- expm1(predict(m, dtest))
  }

  test_mean <- Reduce("+", test_preds_list) / K
  list(oof = oof_preds, test = test_mean)
}

X_train_mat <- as.matrix(X_train)
X_test_mat  <- as.matrix(X_test)

cat("Building OOF predictions for FRE_NGOAITRU ...\n")
res_fre_ng   <- oof_predict_component("FRE_NGOAITRU",   X_train, train_dummy$FRE_NGOAITRU,   X_test_mat)

cat("Building OOF predictions for FRE_NOITRU ...\n")
res_fre_no   <- oof_predict_component("FRE_NOITRU",     X_train, train_dummy$FRE_NOITRU,     X_test_mat)

cat("Building OOF predictions for COST_NGOAITRU ...\n")
res_cost_ng  <- oof_predict_component("COST_NGOAITRU",  X_train, train_dummy$COST_NGOAITRU,  X_test_mat)

cat("Building OOF predictions for COST_NOITRU ...\n")
res_cost_no  <- oof_predict_component("COST_NOITRU",    X_train, train_dummy$COST_NOITRU,    X_test_mat)

# Append OOF predictions as new features
X_train$OOF_FRE_NGOAITRU  <- pmax(res_fre_ng$oof,  0)
X_train$OOF_FRE_NOITRU    <- pmax(res_fre_no$oof,  0)
X_train$OOF_COST_NGOAITRU <- pmax(res_cost_ng$oof, 0)
X_train$OOF_COST_NOITRU   <- pmax(res_cost_no$oof, 0)

X_test$OOF_FRE_NGOAITRU   <- pmax(res_fre_ng$test,  0)
X_test$OOF_FRE_NOITRU     <- pmax(res_fre_no$test,  0)
X_test$OOF_COST_NGOAITRU  <- pmax(res_cost_ng$test, 0)
X_test$OOF_COST_NOITRU    <- pmax(res_cost_no$test, 0)

# Also add log versions of OOF predictions
X_train$LOG_OOF_COST_NGOAITRU <- log1p(X_train$OOF_COST_NGOAITRU)
X_train$LOG_OOF_COST_NOITRU   <- log1p(X_train$OOF_COST_NOITRU)
X_test$LOG_OOF_COST_NGOAITRU  <- log1p(X_test$OOF_COST_NGOAITRU)
X_test$LOG_OOF_COST_NOITRU    <- log1p(X_test$OOF_COST_NOITRU)

cat("OOF stacking features added. New dimensions:\n")
cat("X_train:", nrow(X_train), "×", ncol(X_train), "\n")
cat("X_test :", nrow(X_test),  "×", ncol(X_test),  "\n")
```

---

## 9. Two-Part (Hurdle) Model — Classifier (11.3)

```{r hurdle-classifier}
# Binary target: did this person have ANY cost?
y_binary <- as.integer(y_train > 0)
cat("Proportion with cost > 0:", mean(y_binary), "\n")

ctrl_clf <- trainControl(
  method        = "cv",
  number        = 5,
  verboseIter   = FALSE,
  allowParallel = TRUE,
  classProbs    = TRUE,
  summaryFunction = twoClassSummary
)

y_binary_factor <- factor(ifelse(y_binary == 1, "yes", "no"),
                          levels = c("no", "yes"))

clf_grid <- expand.grid(
  nrounds          = c(300, 600),
  max_depth        = c(4, 6),
  eta              = c(0.05, 0.1),
  gamma            = 1,
  colsample_bytree = 0.8,
  min_child_weight = 10,
  subsample        = 0.8
)

set.seed(42)
model_clf <- train(
  x         = X_train,
  y         = y_binary_factor,
  method    = "xgbTree",
  trControl = ctrl_clf,
  tuneGrid  = clf_grid,
  metric    = "ROC",
  verbosity = 0
)

cat("Best classifier AUC:", max(model_clf$results$ROC), "\n")
print(model_clf$bestTune)

# Probabilities P(cost > 0)
p_positive_train <- predict(model_clf, X_train, type = "prob")[, "yes"]
p_positive_test  <- predict(model_clf, X_test,  type = "prob")[, "yes"]
```

---

## 10. Log-Transform Target & Tune with caret (11.2 + 11.5)

Train the main regressor on `log1p(y)` for positive-cost observations only.

```{r caret-tune}
# ── Filter to non-zero rows for regressor ──────────────────────────────────
pos_idx     <- which(y_train > 0)
X_train_pos <- X_train[pos_idx, ]
y_train_pos <- y_train[pos_idx]
y_train_log <- log1p(y_train_pos)

cat("Training regressor on", length(pos_idx), "positive-cost rows\n")

# ── Cross-validation setup ──────────────────────────────────────────────────
ctrl <- trainControl(
  method      = "cv",
  number      = 5,
  verboseIter = TRUE,
  allowParallel = TRUE
)

# ── Hyperparameter grid ─────────────────────────────────────────────────────
xgb_grid <- expand.grid(
  nrounds          = c(500, 1000),
  max_depth        = c(4, 6),
  eta              = c(0.05, 0.01),
  gamma            = c(1, 2),
  colsample_bytree = c(0.8, 1.0),
  min_child_weight = c(10, 20),
  subsample        = c(0.7, 0.8)
)

cat("Total combinations:", nrow(xgb_grid), "\n")

set.seed(42)
model_caret <- train(
  x         = X_train_pos,
  y         = y_train_log,      # <<< log1p-transformed target
  method    = "xgbTree",
  trControl = ctrl,
  tuneGrid  = xgb_grid,
  metric    = "RMSE",
  verbosity = 0
)

cat("\nBest hyperparameters:\n")
print(model_caret$bestTune)
cat("\nBest CV RMSE (log scale):", min(model_caret$results$RMSE), "\n")

plot(model_caret, main = "CV RMSE by hyperparameter combination (log scale)")
```

---

## 11. Refine `nrounds` with Early Stopping (11.5)

```{r early-stopping}
best_params       <- as.list(model_caret$bestTune)
best_params$nrounds <- NULL   # xgb.cv controls nrounds itself
best_params$objective <- "reg:squarederror"

dtrain_pos <- xgb.DMatrix(as.matrix(X_train_pos), label = y_train_log)

set.seed(42)
cv_result <- xgb.cv(
  params                = best_params,
  data                  = dtrain_pos,
  nrounds               = 3000,
  nfold                 = 5,
  metrics               = "rmse",
  early_stopping_rounds = 50,
  verbose               = 0
)

best_nrounds <- cv_result$best_iteration
cat("Optimal nrounds from early stopping:", best_nrounds, "\n")

# Re-train final model on ALL positive-cost rows with optimal nrounds
final_model <- xgb.train(
  params  = best_params,
  data    = dtrain_pos,
  nrounds = best_nrounds,
  verbose = 0
)

cat("Final model trained with", best_nrounds, "rounds\n")
```

---

## 12. Feature Importance

```{r importance}
imp_mat <- xgb.importance(model = final_model)
xgb.plot.importance(imp_mat, top_n = 20, main = "Top 20 Feature Importance")
```

---

## 13. Predict & Combine Hurdle Model (11.3)

```{r predict-hurdle}
# Regressor predictions (log scale → original scale)
y_reg_pred_test  <- expm1(predict(final_model, xgb.DMatrix(as.matrix(X_test))))

# Final prediction = P(cost > 0) × E[cost | cost > 0]
y_pred_hurdle <- p_positive_test * pmax(y_reg_pred_test, 0)

cat("Prediction summary (hurdle model):\n")
print(summary(y_pred_hurdle))
cat("Proportion predicted as zero (< 0.5 VND):", mean(y_pred_hurdle < 0.5), "\n")
cat("Unique predictions:", length(unique(round(y_pred_hurdle, 3))), "\n")
```

---

## 14. Submit

```{r submit}
submission <- data.frame(
  ID    = test_raw$ID,
  y.hat = pmax(y_pred_hurdle, 0)   # clip negatives — cost cannot be < 0
)

write_csv(submission, "submission.csv")
cat("Saved submission.csv —", nrow(submission), "rows\n")
head(submission)
```
